---
title: "Group 6 R Project"
author: "Beau Barber"
date: "9/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Install tidytuesday package
```{r}
install.packages("tidytuesdayR")
```



#Pull Data

```{r}
rainfall <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-07/rainfall.csv')
temperature <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-07/temperature.csv')
```
## Load all necessary libraries
Be sure to install these packages on your computer if you have never done so before (ex: install.packages("lsmeans"))
```{r}
library(lmerTest)
library(lsmeans)
library(multcomp)
library(multcompView)
library(tidyverse)
library(lawstat)
```

Add a column to rainfall data so the date is all in one column
```{r}
rainfall$date <- paste(rainfall$year,"-",rainfall$month,"-",rainfall$day, sep = "")

```

Filter the rainfall and temperature data so we are only working with variables from Jan 2008 to Dec 2018
```{r}
rainfall_1 <- rainfall %>% filter(date >= "2008-01-01", 
                                 date <= "2018-12-31")
temp_1 <- temperature %>% filter(date >= "2008-01-01", 
                                 date <= "2018-12-31")
head(rainfall_1)
head(temp_1)
```

The ~ variables return the measurements by all combinations of var1 + var2, e.g. var1_1+var2_1, var1_2+var2_1, var1_1+var2_2, var1_2+var2_2. So, for individual variables where you want the measurements across all levels of the others, keep them separate.

```{r}
library(doBy)

summaryBy(temperature ~ temp_type + city_name, data = temp_1,
  FUN = function(x) { c(average = mean(x, na.rm = T), standard_deviation = sd(x, na.rm = T), maximum = max(x, na.rm = T), minimum = min(x, na.rm = T)) } )

summaryBy(rainfall ~ city_name, data = rainfall_1,
  FUN = function(x) { c(average = mean(x, na.rm = T), standard_deviation = sd(x, na.rm = T), maximum = max(x, na.rm = T), minimum = min(x, na.rm = T)) } )

summaryBy(rainfall ~ month, data = rainfall_1,
  FUN = function(x) { c(average = mean(x, na.rm = T), standard_deviation = sd(x, na.rm = T), maximum = max(x, na.rm = T), minimum = min(x, na.rm = T)) } )

summaryBy(rainfall ~ year, data = rainfall_1,
  FUN = function(x) { c(average = mean(x, na.rm = T), standard_deviation = sd(x, na.rm = T), maximum = max(x, na.rm = T), minimum = min(x, na.rm = T)) } )

```


```{r}

```

